MassWateR: Improving quality control, analysis, and sharing of water quality data

### 2023-09-26 manu-draft.qmd
resubmitted

execute: 
  echo: false
  warning: false
  message: false
  ft.keepnext: false
  ft.align: left
  fig-format: svg
  fig-dpi: 300
---

@@

To ensure that *MassWateR* is a known resource for potential users and to encourage its use for QC reporting, analysis, and data submission, a community of practice [e.g., @Meyer22] was established during package development and following its initial release on CRAN (January, 2023). This included several beta testing and training workshops to gather feedback on anticipated data analysis workflows and to educate potential users on appropriate use of the package [@whitehead2007; @list2017]. Many users have not previously been exposed to R for data analysis and a substantial portion of the trainings included an [introduction](https://massbays-tech.github.io/intro-to-r/) to R, as well as *MassWateR*. Emphasis was placed on simple use of the core functions, as opposed to developing custom workflows that combined core R functions with *MassWateR*. However, users were encouraged to learn how to extend the use of the package by leveraging additional R packages, such as *ggplot2* [@Wickham16] to create additional visualizations for modifying the existing analysis plots in *MassWateR*. For example, a [vignette](https://massbays-tech.github.io/MassWateR/articles/modifying.html) was included on the package website to demonstrate how these plots can be modified for custom output using *ggplot2*. Further, a [community forum](https://massbays.discourse.group/c/masswater-r-tools/5) was created as a resource for users to post questions about the package and for others to view the discussion if similar issues were encountered. This approach followed the model used by other popular web forums (e.g., \[StackOverflow\](https://stackoverflow.com/)) for troubleshooting software issues by minimizing duplication of topics through sharing solutions in a public forum.

>>>>

execute: 
  echo: false
  warning: false
  message: false
  ft.keepnext: false
  ft.align: left
  fig-format: tiff
  fig-dpi: 300
---

@@

To ensure that *MassWateR* is a known resource for potential users and to encourage its use for QC reporting, analysis, and data submission, a community of practice [e.g., @Meyer22] was established during package development and following its initial release on CRAN (January, 2023). This included several beta testing and training workshops to gather feedback on anticipated data analysis workflows and to educate potential users on appropriate use of the package [@whitehead2007; @list2017]. Many users have not previously been exposed to R for data analysis and a substantial portion of the trainings included an [introduction](https://massbays-tech.github.io/intro-to-r/) to R, as well as *MassWateR*. Emphasis was placed on simple use of the core functions, as opposed to developing custom workflows that combined core R functions with *MassWateR*. However, users were encouraged to learn how to extend the use of the package by leveraging additional R packages, such as *ggplot2* [@Wickham16] to create additional visualizations for modifying the existing analysis plots in *MassWateR*. For example, a [vignette](https://massbays-tech.github.io/MassWateR/articles/modifying.html) was included on the package website to demonstrate how these plots can be modified for custom output using *ggplot2*. Further, a [community forum](https://massbays.discourse.group/c/masswater-r-tools/5) was created as a resource for users to post questions about the package and for others to view the discussion if similar issues were encountered. This approach followed the model used by other popular web forums (e.g., [StackOverflow](https://stackoverflow.com/)) for troubleshooting software issues by minimizing duplication of topics through sharing solutions in a public forum.
